\chapter*{Preface}


This is the end of a long journey to complete my Master's of Science in Civil Engineering, with a specialization in Coastal Engineering. It has been 2 years of work under sometimes less-than-ideal conditions caused by a global pandemic. But I am lucky to have written this thesis at the TU Delft and with Van Oord. Moving to a foreign country during a global pandemic sometimes feel almost reckless in retrospect, but I have grown and learned so much from all the fantastic people I have met here at the TU that I'm incredibly glad I made the leap.  



I would like to thank my committee for their input and guidance throughout this project. Especially Gerben de Boer, for choosing me to work with him and his team at the Datalab, and for some very nerdy conversations about our shared love of Docker. Your enthusiasm, sharp questions, and constant ideas were integral to the final product. Thank you for contributing your time and your team's resources to my project. Thanks to Alejandra Gijón Mancheño, for your help shaping the early version of my project, and your flexibility as the scope changed. Your feedback, and our conversations about how to approach things really helped me develop as a student/researcher. To Bert Wouters, thank you for your feedback on my writing and providing some answers to my ICESat-2 questions. And finally, thank you to my chair Stefan Aarninkhof for your encouragement and enthusiasm during committee meetings, and your suggestions which also gave the project shape.



Finally, thank you to my parents and my entire family for their unending support --- emotional, financial and otherwise. None of this would have been possible without your love and support. Finally thank you to my friends: new friends here in Delft who kept me sane in the darkest days of the pandemic, and my longtime friends on the East coast, and all over the world who I still am close to, despite living on another continent. Thank you so much, for everything.

\begin{flushright}
{\makeatletter\itshape
    \@author \\
    Delft, October 2022
\makeatother}
\end{flushright}

